---
title       : Investigating Variable Effects on MPG
subtitle    : Utilizing Parallel Coordinate Graphs
author      : Steve Rowe
job         : 
framework   : revealjs     # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {selfcontained,  standalone, draft}
knit        : slidify::knit2slides

---

# Investigating Variable Effects on MPG

## Utilizing Parallel Coordinate Graphs


### By: Steve Rowe

--- .class #id 

## Parallel Coordinate Graphs

* Way of visualizing multivariate data
* Used as early as 1880 by Henry Gannetts in *General Summary, Showing the Rank of States, by Ratios*
* More useful when used interactively, as in this [application](https://thornkin.shinyapps.io/dataproducts):

``` {r, echo=F, fig.height=5}
library(MASS)
library(RColorBrewer)
parcoord(mtcars,col=brewer.pal(12, "Paired"), var.label=T)
```

---

## Using Parallel Coordinates With MPG
The [application](https://thornkin.shinyapps.io/dataproducts) allows two axes of control:
* Selecting Variables
 * Used to see relationships more clearly
* Limiting MPG range
  * Used to see which variables are related to ranges.
  * For instance, what kind of cylinders are in high MPG cars?

---
## Selecting Variables
* Variables can be removed by unchecking the variables boxes.
 * Sometimes relationships can become more clear by limiting the variables.
* For example, only displacement and weight looks like this:

``` {r, echo=F, fig.height=3, warning=F}
library(dplyr)
cars <- select(mtcars, mpg, disp, wt)
parcoord(cars,col=brewer.pal(12, "Paired"), var.label=T)
```
* It is clear that displacement is inversely correlated with high MPG.  Weight is highly correlated with displacement.

---

## Utilizing the MPG Range
* The MPG range can be limited to see what variables affect a high MPG cars.
* Limiting the range to >= 22 MPG generates this graph:
``` {r, echo=F, fig.height=4, warning=F}
cars <- filter(mtcars, mpg > 22) %>% select(-cyl, -vs, -gear, -carb)
parcoord(cars,col=brewer.pal(12, "Paired"), var.label=T)
```
* Most high MPG cars have low displacement, low weight, are slow, and are manual transmissions.
